Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-14 WSSSPE1_report/WSSSPE1_summary_report.tex
commenting a note
Adding more papers from Karen's summary.

@@

\note{More papers are needed here.}

@@

Summary of above discussed cases: lessons, trends, commonalities, is there an
ideal sustainability scenario?

>>>>

Drawing on experiences from high-energy physics, \cite{Vay_WSSSPE} proposed
developing teams of technical specialists able to overcome a lack of
coordination between projects.  Maximization of scientific output requires
maximizing the usability of the scientific codes while minimizing the cost of
developing and supporting those codes.  This included targeting different
architectures for their software to be deployed, as well as coordination
between technically-focused individuals and usage of a common scripting
language between projects.  Instead of fragmenting the development of
simulation codes across institutions, the paper suggests that a cohesive
strategy reducing duplication and increasing coordination will broadly increase
the efficiency across institutions.  The approach proposed is of de-fragmenting
the existing ecosystem in a non-disruptive way.

@@

In \cite{Hanwell_WSSSPE}, the authors reflect on the 15-year history of open
source software development at Kitware.  In particular, they focus on their
success at growing their community of users through enabling multiple channels
of communication, directly reaching out to individuals, and lowering the
barrier to entry for contributions.  This involves providing clear,
test-oriented and review-based mechanisms for evaluating contributions,
permissive licenses, and a service-based model for sustaining development.
This model enables Kitware to review both public funding, as well as private
funding to support improvements and targeted developments of the software.

At NESCent, a combination of in-house informatics individuals and domain
scientists collaborate to develop software to study evolutionary science.  The
report, \cite{Cranston_WSSSPE}, studied the success of a ``hackathon'' model
for development, where short-form, hands-on events combining users,
researcher-developers and software engineers targeted specific code
improvements.  From this experiment, the authors identified sweveral key
outcomes as well as lessons-learned -- specifically, the co-localization of
developers was seen as having a strong impact, enabling casual conversation
that led to discrete outcomes.  The formation of the discussion mailing list,
specifically in response to the social capital built at the hackathon, was seen
as spurring on longer-term benefits to the community and fostering
sustainability.

\cite{Hart_WSSSPE} addresses the success of the rOpenSci project in developing
collaborations supporting tool development for Open Science.  This software
collective, organized around the statistical programming environment R,
develops access mechanisms for data repositories and attempts to reduce the
barrier to entry for individuals wanting to access data repositories and study
the data contained therein.  The collective fosters direct collaboration
between individuals and data providers, designed to ``train academics in
reproducible science workflows focused around R.''  The two central challenges
to this goal were identified as engagement of existing users within ecology and
evolutionary biology (EEB), and how can the community make inroads and traction
in other disciplines?  Currently, the collective is exploring addressing these
challenges through use of social media and holding workshops and hackathons.
This helps to both raise the profile of the collective within EEB and in other
domains.  However, the overarching challenge identified in the paper was that
of incentivizing maintenance of software, which is difficult in academia.

@@

%Summary of above discussed cases: lessons, trends, commonalities, is there an ideal sustainability scenario?
